MLExplain

### 2018-05-28 paper.tex
talk about record patterns
final tweaks?

At first glance, JSExplain seems tightly linked to JavaScript. However, the
architecture of the project is quite modular, and can be used for a different
programming language, by providing an interpreter written in the subset of OCaml
that we support as well as a way to display OCaml values.

@@

To interpret OCaml code, we first need to parse and type it. However, our
compiler does not allow us to use the standard library nor an external one. As
we do not need to trace the parsing nor the typing, we have been able to
separate totally the frontend from the interpreter itself and use another
compiler to compile it. Figure \ref{fig:archi} describes the architecture of
the whole application.

@@

The use of \texttt{ppx} lets us write the much simpler code, which is
transformed to the previous version during parsing.
\begin{verbatim}
let%some a = function_that_could_fail 0 in
continuation a
\end{verbatim}

@@

We give some statistics about the project in Figure~\ref{fig:statistics}. As can
be seen, the largest part of the code is the frontend, in particular the
conversion from OCaml's AST to our format. The interpreter is quite short,
because the execution of OCaml is actually very simple. Note that we support all
of OCaml, with the exception of objects.

@@

\section{Conclusion}
We have shown how JSExplain can be easily adapted to other programming language,
in this case OCaml. Most of the complexity of OCaml being its typing, writing
the interpreter was quite straightforward. As future work, we plan to extend the
subset of OCaml supported by our compiler, to be able to directly use the
standard library when writing interpreters.

>>>>

At first glance, JSExplain seems tightly linked to JavaScript. However, the
architecture of the project is quite modular and can be used for a different
programming language, by providing both an interpreter written in the subset of
OCaml that we support and a way to display OCaml values.

@@

To interpret OCaml code, we first need to parse and type it. However, our
compiler does not allow us to use the standard library nor an external one, as
they are not written in the subset of OCaml that we support. As we do not need
to trace the parsing nor the typing, we have been able to separate totally the
frontend from the interpreter itself and use another compiler to compile it.
Figure \ref{fig:archi} describes the architecture of the whole application.

@@

The use of \texttt{ppx} lets us write much simpler code, which is transformed to
the previous version during parsing.
\begin{verbatim}
let%some a = function_that_could_fail 0 in
continuation a
\end{verbatim}

@@

We give some statistics about the project in Figure~\ref{fig:statistics}. As can
be seen, the largest part of the code is the frontend, in particular the
conversion from OCaml's AST to our format. The interpreter is quite short,
because the execution of OCaml is actually very simple. Note that we support all
of OCaml, with the exception of objects and record patterns.

@@

\section{Conclusion}
We have shown how JSExplain can be easily adapted to other programming language,
in this case OCaml. Most of the complexity of OCaml being its typing, writing
the interpreter was quite straightforward. The current implementation of our
interpreter is very simple and contains a lot of documentation. The architecture
of the code is as unsurprising as possible, enabling easy extensions. As the
design of our interpreter is very simple, it could be considered as a first step
towards a specification of OCaml. MLExplain also provides a comfortable
environment to experiment with the semantics of OCaml. It is thus a good
playground for future evolution of the language.

As future work, we plan to extend the subset of OCaml supported by our compiler,
to be able to directly use the standard library when writing interpreters, and
to implement the few OCaml construct missing from our interpreter. We also want
to improve the web interface and to allow the user to load code from multiple files.
